---
title: "Drosophila Traumatic Brain Injury"
author: "Anji Trujillo - Professor Coon"
date: "December 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
require(knitr)
opts_knit$set(root.dir = "E:/Projects/Proteomics/")
knitr::opts_chunk$set(root.dir = "E:/Projects/Proteomics/", warning = FALSE, message = FALSE)
```

This is an R Markdown Document describing the proteomics data collected for the Drosophila Traumatic Brain Injury Project in collaboration with Professor Wassarman and Becky Steinbrech. The samples analyzed were derived from heads and hemolymph of Drosophilla melanogaster. For the head samples, there are 17 time points with a control and a traumatic brain injury (TBI) sample at each time point. The hemolymph sample is a shorter temporal study with control and traumatic brain injury samples for 4 time points.

Proteomics data was collected September 22, 2018 and searched on October 16, 2018. The initial analysis was performed December 9, 2018.

The head and hemolymph data was searched separately with an appropriate fractionated sample to increase our protein group resolution. 

```{r install_packages, echo=FALSE}
library(readr)
library(plyr)
library(dplyr)
library(pheatmap)
library(pcaMethods)
library(ggplot2)
library(devtools)
library(e1071)
library(dplyr)
library(Mfuzz)
library(cluster)
library(yaml)
library(Rcpp)
```

## Load data
```{r load_data, echo=FALSE}
proteinGroups_dros_hemo <- read_delim("E:/Projects/Proteomics/DorsophilaHead_Experiment/txt_hemo_plusFrac/proteinGroups.txt","\t", escape_double = FALSE, trim_ws = TRUE)
cat(paste0("Number of protein groups in Hemolymph Data before any filtering of missing measurements: ", nrow(proteinGroups_dros_hemo))) #, "\n", "Number of columns: ", ncol(proteinGroups_dros_hemo)))
#proteinGroups_dros_hemo <- data.frame(proteinGroups_dros_hemo)
proteinGroups_dros_heads <- read_delim("E:/Projects/Proteomics/DorsophilaHead_Experiment/txt_dros_heads_plusFrac/proteinGroups__dros_heads_plusFrac.txt","\t", escape_double = FALSE, trim_ws = TRUE)
cat(paste0("Number of protein groups in Heads Data before any filtering of missing measurements: ", nrow(proteinGroups_dros_heads))) #, "\n", "Number of columns: ", ncol(proteinGroups_dros_heads)))

```

The data analysis was performed using R. The following functions were written to manipulate the data:  
1. Remove Contaminants and Reverse Sequences  
2. Subset columns only with protein ID's, number identifier, and Gene Name  
3. 50% data cut off, filters data to exclude proteins that have more than 50% of measurements missing 
4. Plotting Standard Deviation of Proteins
5. Fuzzy c-means preperation for data that has not been imputed yet  
6. Fuzzy c-means for data that has been previously imputed (in this case Bayesian Statistics from PCAMethods package)  

```{r functions, echo=FALSE}
RemoveContaminants <- function(x){
 x <- x[-grep("*",x$`Potential contaminant`),]
 x <- x[-grep("*",x$Reverse),]
 x <- x[-grep("*",x$`Only identified by site`),]
 x <- x[-grep("CON",x$`Protein IDs`),]
 x <- x[-grep("REV",x$`Protein IDs`),]
 x <- x[,-which(names(x) %in% c("Potential contaminant","Reverse","Only identified by site"))]
 return(x)
} 

subsetLFQ <- function(x){
  y <- x[,which(names(x) %in% c("Protein IDs", "id","Gene names"))]
  z <- x[,grep("LFQ intensity",names(x))]
  z[z == 0] <- NA
  x <- data.frame(y,z)
  return(x)
}
 
remove.features.50percentcuttoff <- function (x) {
  features.missing = rowMeans(is.na(x)) 
  print(paste0("Number of protein groups that have over 50% missing measurements: ",sum(features.missing > 0.50))) 
  features.missing.50more = rownames(x)[features.missing > 0.50] 
  
  keep.features = which(features.missing <= 0.50) 
  print(paste0("Protein groups that pass the 50% filteration: ", length(keep.features)))
  names(keep.features) = keep.features 
  
  remove.features = which(features.missing > 0.50)
  print(paste0("Number of protein groups removed from dataset: ", length(remove.features)))
  names(remove.features) = remove.features
  
  filtered = x[-which(rownames(x) %in% remove.features),]
  return(filtered)
}

filter.std.plotting <- function (eset, min.std,visu=TRUE)
{
  #index <- logical(dim(exprs(eset))[1])

  tmp <- logical(dim(exprs(eset))[1])
  if (is.numeric(min.std))
  { 
    data <- exprs(eset)
      for (i in 1:length(tmp))
      {
        tmp[i]   <- sd(data[i,],na.rm=TRUE)
        #index[i]  <- ( tmp[i] > min.std)
        
      }
    index <- tmp > min.std
    index[is.na(index)] <- TRUE
    cat(paste(sum(!index),"Proteins have a standard deviation greater than ", min.std, ".\n"))
  }
  
  if (visu)
  {
    plot(sort(tmp),xlab="Ordered Hemo Proteins",ylab="Standard Deviation")
  }
  eset[index,]
}

fuzzyprep_imputation_included <- function(z)
{
  exprValues <- new("ExpressionSet", exprs = as.matrix(z))
  # exclude proteins that have more than 50% of measurements missing
  exprValues.r <- filter.NA(exprValues, thres = 0.50)
  # Fuzzy c-means does not allow for missing values, replace missing values by median values
  exprValues.f = fill.NA(exprValues.r, mode="median")
  # Set a minimum threshold for variation 
  tmp = filter.std(exprValues.f, min.std = 0.1)
  # Clustering is performed in Eculidian space, standaridize abundance values to have a mean value of zero
  # Ensures that proteins with similar changes in abundance are close in Euclidean space
  exprValues.s = standardise(tmp)
  return(exprValues.s)
}

fuzzyprep_usepreviousImputation <- function(z)
{
  exprValues <- new("ExpressionSet", exprs = as.matrix(z))
  tmp = filter.std.plotting(exprValues, min.std = 0.1)
  exprValues.s = standardise(exprValues)
  return(exprValues.s)
}

fuzzyprep_usepreviousImputation_foldchange <- function(z)
{
  exprValues <- new("ExpressionSet", exprs = as.matrix(z))
  tmp = filter.std.plotting(exprValues, min.std = 0.1)
  #exprValues.s = standardise(exprValues)
  return(exprValues)
}

```
## Hemolymph Data
The hemolymph data includes samples for the first 24 hours (1,4,8,24 hrs) after traumatic brain injury has been implemented. The data analysis will:  
1. Clean proteomics data from  MaxQuant search algorithm output  
2. Distribution of proteomic data before and after Bayesian imputation.  
3. PCA plots describing the variation in protein groups (scores) and the variation in the samples (loading) 
4. Fuzzy Clustering and hierarchical plots  
5. Clustering is also performed on fold change data


```{r subset_hemo, include=FALSE}
proteinGroups_dros_hemo <- RemoveContaminants(proteinGroups_dros_hemo)
print(dim(proteinGroups_dros_hemo))
proteinGroups_dros_hemo <- subsetLFQ(proteinGroups_dros_hemo) 
print(dim(proteinGroups_dros_hemo))
```

Number of missing measurement in Hemo dataset: `r sum(is.na(proteinGroups_dros_hemo))`

Filter and exclude protein groups that are missing over 50% of measurements 
```{r filter_hemo, echo=FALSE}
filtered_dros_hemo_50percent <- remove.features.50percentcuttoff(proteinGroups_dros_hemo)
```

Implement a bayesian pca imputation for data that is log2 transformed. The R package used is called [pcaMethod](https://www.bioconductor.org/packages/devel/bioc/manuals/pcaMethods/man/pcaMethods.pdf).
```{r bayesian_setup, echo=FALSE}
#subset matrix containing only samples that have been filtered 50% threshold
hemo_50percent <- as.matrix(filtered_dros_hemo_50percent[,c(4:11)])
#log2 transform
hemo_50percent_log2 <- log2(hemo_50percent)
#set row names to match protein group identifer number
rownames(hemo_50percent_log2) <- filtered_dros_hemo_50percent$id
colnames(hemo_50percent_log2) <- c("Ctr_1hr", "1hr", "Ctr_4hr","4hr", "Ctr_8hr", "8hr", "Ctr_24hr", "24hr")

hemo_meta <- data.frame(HemoTimePoints = colnames(hemo_50percent_log2),Sample_Type=rep(c("Control","TBI"),4))
```



```{r bayesian, include=TRUE}
#PCA function utelizing method=bpca "bayesian"
pc_hemolog2 <- pca(hemo_50percent_log2, nPcs = 3, method = "bpca") #pca method
#extract imputed data set for hemo data
imputed_hemo <- completeObs(pc_hemolog2)
```

```{r write_csv_imputedhemo, echo=FALSE}
subset <- proteinGroups_dros_hemo[,1:2]
subset = subset[which(rownames(subset) %in% rownames(imputed_hemo)), ]
imputed_hemo_subset <- data.frame(subset,imputed_hemo)
colnames(imputed_hemo_subset) <- c("Protein.ID" , "Gene.ID", "Ctr_1hr", "1hr", "Ctr_4hr","4hr", "Ctr_8hr", "8hr", "Ctr_24hr", "24hr")
write.csv(imputed_hemo_subset, "E:/Projects/Proteomics/DorsophilaHead_Experiment/Imputed_hemo_log2.csv")
```


The distribution of the data bfore and after Bayesian imputation, plotted  below.
```{r density_hemo, echo=FALSE}
hist(hemo_50percent_log2, breaks = 20, xlab = "Log2(Hemo data with the 50% filter)", main = "Histogram of Hemo Data Before Imputation")

```

```{r even_odd, echo=FALSE}
#Split data up between the controls and the TBI samples
even_index <- seq(2,8,2) #TBI samples
odd_index <- seq(1,8,2) #controls

#TBI Samples
imputed_hemo_even <- imputed_hemo[,even_index]
#Control samples
imputed_hemo_odd <- imputed_hemo[,odd_index]
#Reorder data so that first 4 samples are Controls and second set of 4 are the TBI samples
imputed_hemo_reorder <- imputed_hemo[,c(odd_index,even_index)]

```

```{r density_imputed_hemo, echo=FALSE}
hist(imputed_hemo_reorder, breaks = 20, xlab = "Log2(Imputed Hemo Data with 50% filter", main = "Histogram of Imputed Hemo Data")
```

A global view of the data can visualized by PCA analysis. Imputed data is used from here on forward     
* The first PCA is a plot where each dot represents a protein (6,030 proteins), the text denote which sample the protein group derives from, and the color indicates either control or TBI sample type.  
* In the second PCA plot the loadings describe the variance in each of the hemolymph samples determined by protein group. Each dot is a hymolymph sample and the color indicates either control or TBI sample type.  
```{r pca_plot, echo=FALSE}
palette(c("mediumorchid2","mediumturquoise","olivedrab3", "darkgoldenrod1", 
          "hotpink3", "red2", "steelblue2", "sienna2","slategray4", 
          "deepskyblue", "orangered", "midnightblue"))

# PCA Plot for Scores of Log2(hemo data)
sample.colors = as.numeric(factor(hemo_meta$Sample_Type))
plot(scores(pc_hemolog2), pch = 16, col = sample.colors, main = "PCAScores Hemolymph Log2(Protein)")
text(scores(pc_hemolog2)[,1], scores(pc_hemolog2)[,2], labels = colnames(hemo_50percent_log2), 
       col = sample.colors)
  legend("bottomright", legend = levels(hemo_meta$Sample_Type), pch = 16, col = 1:length(levels(hemo_meta$Sample_Type)),
         y.intersp = 0.7)
  
# PCA Plot the loadings of Log2(hemo data)
plot(loadings(pc_hemolog2), pch = 19, col = sample.colors, main = "PCALoadings") 
text(loadings(pc_hemolog2)[,1], loadings(pc_hemolog2)[,2], labels = colnames(hemo_50percent_log2), 
       col = sample.colors)
  legend("bottomleft", legend = levels(hemo_meta$Sample_Type), pch = 16, col = 1:length(levels(hemo_meta$Sample_Type)),
         y.intersp = 0.7)

```

  
## Hemo mFuzz Soft Clustering
Clustering will be implemnted on the proteomics data set that that has been organized such that control samples are the first and TBI samples are following.

Prior to fuzzy clustering the variation for each protein plotted. 
```{r mfuzz_bayesian_imputed, echo=FALSE}

z = imputed_hemo_reorder # change accordingly to use in mfuzzy setup funtions

exprValues.s_bayesian <- fuzzyprep_usepreviousImputation(z)
```



Soft clustering is implemented in the function mfuzz using a fuzzy c-means algorithm from [e1071](https://cran.r-project.org/web/packages/e1071/e1071.pdf) package.
```{r mfuzz_bayesian, echo=FALSE}
error <- NA #set error to NA

y = exprValues.s_bayesian #change according to version of data to perform fuzzy clustering
```

Fuzzy clustering is one many types of clustering algoriths to choose from e.g. K-means, hierarchical, bayesian, etc. I will show fuzzy clustering along with hierarchical clustering to visualize the protein trends we see for through out the timepoint. The clustering will be performed first on data set that includes both the control and TBI samples and then performed on the fold change data relative to control.

To begin th we must first optimize the parameters for fuzzy clustering we must calculate the fuzzier value as well as the number of clusters to divide the data into. The fuzzier "m" and the number of clusters "c" must be chosen in advanced. This fuzzy type of clustering is advantageous over hard cluster (e.g. k-means) which commonly detects clusters of random data. 

The function mestimate 
```{r mfuzz_fuzzier, echo=FALSE}
m1 <- mestimate(y) 
print(paste0("fuzzier m = ",m1))
```

```{r mfuzz_numberOfClusters, include=TRUE}
# Find the error associated with number of cluster 2-15
for(i in 2:15){
  c1 <- mfuzz(y, c=i, m=m1)
  error <- rbind(error, c(i,c1$withinerror))
}
```

Look at the error calculated for each of the clusters.
```{r Error_plot, echo=FALSE}
plot(error[,1], error[,2], xlab = "Cluster Number", ylab = "Error")
```

Fuzzy clusters are plotted for 6 clusters.   
*Note first 4 samples are Controls and next set of 4 are TBI samples  
*Note a seed has been set to keep clusters consistent.   
```{r numberOfClusters_beforeMembership, echo=FALSE}
c1 <- mfuzz(y, c=6, m=m1)
set.seed(123) #set seed to keep clusters consistent
mfuzz.plot2(y, cl=c1, mfrow = c(3,2), time.labels = c("1","4", "8", "24","1","4", "8", "24"), 
            ylab = "Protein Changes", xlab = "Hemo Time Points (hour)",  col.lab="black", x11=F)
```

Filter out proteins from clusters with a membership score of greater than 70%. Accounting for the majority of the temporal trend.
```{r numberOfClusters_70percentMembership, echo=FALSE}

mfuzz.plot2(y, cl=c1, mfrow = c(3,2), min.mem = 0.70, time.labels = c("1","", "8", "24","1","4", "8", "24"),
            ylab = "Protein Changes", xlab = "Hemo Time Points (hour)",  col.lab="black", x11=F)

```


How many proteins have a membership score greater than 70%, count TRUE.
```{r membershipGreater70_reorder, echo=FALSE}
table(rowSums(c1$membership > .70) >0)
```

Heat map of hemo samples that have a membership value greater than 70%.
Global view of hemo samples.
```{r heatmap_Highmembership, echo=FALSE}
scaleRYG <- colorRampPalette(c("red","black","darkgreen"), space = "rgb")(31)


pheatmap(
  mat               = imputed_hemo_reorder[rowSums(c1$membership >0.7)>0,],
  color             = scaleRYG,
  border_color      = NA,
  scale = "row",
  cluster_cols = FALSE,
  cluster_rows = TRUE,
  clustering_distance_rows = "euclidean",
  labels_col = c("Ctr_1hr","Ctr_4hr", "Ctr_8hr", "Ctr_24hr","1hr","4hr", "8hr", "24hr"),
  show_colnames     = TRUE,
  show_rownames     = FALSE,
  #cellwidth = 40,
  #cellheight = .25,
  #annotation_col    = mat_col,
  #annotation_colors = mat_colors,
  drop_levels       = TRUE,
  fontsize          = 8,
  main              = "Log2 Hemolymph Data with membership scores greater than 70% (scaled by row)"
)

```


```{r distance_reordered,out.width='\\textwidth', fig.height = 8, fig.align='center', include=FALSE}

reorder_hclust <- hclust(dist(imputed_hemo_reorder[rowSums(c1$membership >0.7)>0,], method = "euclidean"))

clusters <- cutree(reorder_hclust, k = 4)

plot(reorder_hclust, label= FALSE)

#rect.hclust(reorder_hclust, k=4, border = "red")
```

```{r clusters_reordered, include=FALSE}
#clusplot(imputed_hemo_reorder, clusters, lines = 0)
```

## Hemo mFuzz Soft Clustering - fold change

Fuzzy clustering is performed on fold change data set. The fold change was calculated relative to controls. We will look at clustering of proteins that have a fold change greater than 1.2. There are fold changes in the hemolymph data that are greater than 8, and these proteins with extreme fold changes must be looked at in depth. However to reduce their bias on clustering, I have implemented a binning strategy that takes a look at the data set between logfold -2 and 2, compressing the larger fold changes down to fold change of 2.1. This way we can see how the protein groups bin at a more granular level. 

I have added fuzzy clustering at the end of this section, however I dont believe fuzzy clustering will be the best strategy for visualizing the clusters. There are hierarchical clusters and a clusplot at the beginning of this section, to help determine the number of clusters the data partitions into.
```{r mfuzz_foldchange, include=TRUE}
imputed_hemo_fold <- imputed_hemo_reorder[,5:8]-imputed_hemo_reorder[,1:4]

write.csv(imputed_hemo_fold, "E:/Projects/Proteomics/DorsophilaHead_Experiment/Imputed_hemo_foldchange.csv")

# subset features with fold change greater than 1.2 to cluster
important_features_hemo_foldchange <- imputed_hemo_fold[rowSums(abs(imputed_hemo_fold)>1.2)>0,]

# bin protein groups so that we describe the changes within fold changes of -2 and 2
important_features_foldchange_compress <- important_features_hemo_foldchange
important_features_foldchange_compress[important_features_foldchange_compress < -2] = -2.1
important_features_foldchange_compress[important_features_foldchange_compress > 2] = 2.1

# Protein groups with fold change greater than 1.2
table(rowSums(abs(imputed_hemo_fold)>1.2)>0)

# Protein groups with fold change greater than 2
table(rowSums(abs(imputed_hemo_fold)>2)>0)

# Protein groups with fold change greater than 4
table(rowSums(abs(imputed_hemo_fold)>4)>0)

# Protein groups with fold change greater than 8
table(rowSums(abs(imputed_hemo_fold)>8)>0)

```

```{r write_csv_imputedhemo_foldChange, echo=FALSE}
imputed_hemofoldChange_subset <- data.frame(subset,imputed_hemo_fold)
colnames(imputed_hemofoldChange_subset) <- c("ProtienGroup.ID", "Gene.ID", "Hemo_1hr","Hemo_4hr", "Hemo_8hr", "Hemo_24hr")
write.csv(imputed_hemo_subset, "E:/Projects/Proteomics/DorsophilaHead_Experiment/Imputed_hemo_foldchange.csv")
```

```{r heatmap_important_features_1.2, echo=FALSE}
pheatmap(
  mat               = important_features_hemo_foldchange, #[rowSums(c2$membership >0.5)>0,],
  color             = scaleRYG,
  border_color      = NA,
  #scale = "row",
  cluster_cols = FALSE,
  cluster_rows = TRUE,
  labels_col = c("1hr","4hr", "8hr", "24hr"),
  #breaks = breaksList,
  clustering_distance_rows = "euclidean",
  show_colnames     = TRUE,
  show_rownames     = FALSE,
  #cellwidth = 40,
  #cellheight = .25,
  #annotation_col    = mat_col,
  #annotation_colors = mat_colors,
  drop_levels       = TRUE,
  fontsize          = 10,
  main              = "Heat map of Hemo protiens with fold change < -1.2, > 1.2"
)

pheatmap(
  mat               = important_features_foldchange_compress, #[rowSums(c2$membership >0.5)>0,],
  color             = scaleRYG,
  border_color      = NA,
  #scale = "row",
  cluster_cols = FALSE,
  cluster_rows = TRUE,
  labels_col = c("1hr","4hr", "8hr", "24hr"),
  #breaks = breaksList,
  clustering_distance_rows = "euclidean",
  show_colnames     = TRUE,
  show_rownames     = FALSE,
  #cellwidth = 40,
  #cellheight = .25,
  #annotation_col    = mat_col,
  #annotation_colors = mat_colors,
  drop_levels       = TRUE,
  fontsize          = 10,
  main              = "Fold Change with compressed Hemo Data, binning <-2 and >2"
)


```

```{r mfuzz_clustering, include=TRUE}
fold_hclust <- hclust(dist(important_features_foldchange_compress, method = "euclidean"))
clusters <- cutree(fold_hclust, k = 6) #k = 20)

clusplot(important_features_foldchange_compress, clusters, lines = 0)

plot(fold_hclust, label= FALSE)
rect.hclust(fold_hclust, k=6, border = "red")
```

# Fold Change Hemo Data Fuzzy Clustering
```{r mfuzz_bayesian_imputed_foldchange, include=TRUE}

#start with the re-ordered hemo data
z = important_features_foldchange_compress # change accordingly to use in mfuzzy setup funtions

exprValues.s_bayesian <- fuzzyprep_usepreviousImputation_foldchange(z)
```

```{r mfuzz_bayesian_foldchange, include=TRUE}
error <- NA #set error to NA

y = exprValues.s_bayesian #change according to version of data to perform fuzzy clustering
```

The function mestimate 
 
```{r mfuzz_bayesian_imputed_foldchange_error, include=TRUE}
m1 <- mestimate(y) 

# 
for(i in 2:15){
  c1 <- mfuzz(y, c=i, m=m1)
  error <- rbind(error, c(i,c1$withinerror))
}
```

Plot error calculated from imputed hemo data set that has been reordered. 
```{r error_hemo_foldChange, echo=FALSE}
plot(error[,1], error[,2])
```

```{r numberOfClusters_foldchange, echo=FALSE}
c1 <- mfuzz(y, c=6, m=m1)
mfuzz.plot2(y, cl=c1, mfrow = c(3,2), time.labels = c("1hr","4hr", "8hr", "24hr"), 
            ylab = "Protein Changes", xlab = "Hemo Time Points",  col.lab="black", x11=F)
mfuzz.plot2(y, cl=c1, mfrow = c(3,2), time.labels = c("1hr","4hr", "8hr", "24hr"), 
            min.mem = 0.50, ylim.set = c(-2,2), ylab = "Protein Changes", xlab = "Hemo Time Points",  col.lab="black", x11=F)

```


## Head Data

```{r subset_head, include=TRUE}
proteinGroups_dros_heads <- RemoveContaminants(proteinGroups_dros_heads)
print(dim(proteinGroups_dros_heads))
proteinGroups_dros_heads <- subsetLFQ(proteinGroups_dros_heads) 
print(dim(proteinGroups_dros_heads))

```

Number of missing measurement in Head dataset: `r sum(is.na(proteinGroups_dros_heads))`

Filter and exclude protein groups that are missing over 50% of measurements 
```{r filter_head, echo=FALSE}
filtered_dros_head_50percent <- remove.features.50percentcuttoff(proteinGroups_dros_heads)
```

Implement a bayesian pca imputation for data that is log2 transformed. The R package used is called [pcaMethod](https://www.bioconductor.org/packages/devel/bioc/manuals/pcaMethods/man/pcaMethods.pdf).
```{r bayesian_setup_head, include=TRUE}
#subset matrix containing only samples that have been filtered 50% threshold
head_50percent <- as.matrix(filtered_dros_head_50percent[,c(5:38)])
#log2 transform
head_50percent_log2 <- log2(head_50percent)
#set row names to match protein group identifer number
rownames(head_50percent_log2) <- filtered_dros_head_50percent$id

colnames(head_50percent_log2) <- as.numeric(gsub(".*_","",colnames(head_50percent_log2)))

col_levels <- levels(colnames(head_50percent_log2)) <- seq(1:34)
colnames(head_50percent_log2) <- order(colnames(head_50percent_log2))

col_names <- names(colnames(head_50percent_log2)) <- order(colnames(head_50percent_log2))

head_meta <- data.frame(HeadTimePoints = order(col_levels),Sample_Type=rep(c("Control","TBI"),17))

head_50percent_log2 <- head_50percent_log2[,names(colnames(head_50percent_log2))]

```

```{r bayesian_heads, include=TRUE}
#PCA function utelizing method=bpca "bayesian"
pc_headlog2 <- pca(head_50percent_log2, nPcs = 3, method = "bpca") #pca method
#extract imputed data set for hemo data
imputed_head <- completeObs(pc_headlog2)

```

```{r write_csv_imputedhead, echo=FALSE}
rownames(proteinGroups_dros_heads) <- proteinGroups_dros_heads$id
subset <- proteinGroups_dros_heads[,1:2]
subset = subset[which(rownames(subset) %in% rownames(imputed_head)), ]
imputed_head_subset <- data.frame(subset,imputed_head)

colnames(imputed_head_subset) <- c("ProtienGroup.ID", "Gene.ID","0","Ctr_.5 hr", "Ctr_1 hr", "Ctr_2 hr","Ctr_4 hr","Ctr_6 hr", "Ctr_8 hr", "Ctr_12 hr", "Ctr_16 hr", "Ctr_24 hr", "Ctr_48 hr", "Ctr_72 hr", "Ctr_168 hr", "Ctr_336 hr", "Ctr_504 hr", "Ctr_672 hr", "Ctr_840 hr", "0",".5 hr", "1 hr", "2 hr","4 hr","6 hr", "8 hr", "12 hr", "16 hr", "24 hr", "48 hr", "72 hr", "168 hr", "336 hr", "504 hr", "672 hr", "840 hr")

write.csv(imputed_head_subset, "E:/Projects/Proteomics/DorsophilaHead_Experiment/Imputed_head.csv")
```

A global view of the data can visualized by PCA analysis. Imputed data is used from here.     
* The first PCA is a plot where each dot represents a protein (7,175 proteins), the text denote which timepoint the protein group derives from, and the color indicates either control or TBI sample type.  
* In the second PCA plot the loadings describe the variance in each of the brain samples determined by protein group. Each dot is a drosphila sample and the color indicates either control or TBI sample type. 
```{r pca_plot_head, echo=FALSE}
palette(c("mediumorchid2","mediumturquoise","olivedrab3", "darkgoldenrod1", 
          "hotpink3", "red2", "steelblue2", "sienna2","slategray4", 
          "deepskyblue", "orangered", "midnightblue"))

# PCA Plot for Scores of Log2(hemo data)
sample.colors = as.numeric(factor(head_meta$Sample_Type))
plot(scores(pc_headlog2), pch = 16, col = sample.colors, main = "PCAScores Head Protein Groups Log 2 Raw Values")
text(scores(pc_headlog2)[,1], scores(pc_headlog2)[,2], labels = colnames(head_50percent_log2), 
       col = sample.colors)
  legend("bottomleft", legend = levels(head_meta$Sample_Type), pch = 16, col = 1:length(levels(head_meta$Sample_Type)),
         y.intersp = 0.7)
  
# PCA Plot the loadings of Log2(head data)
plot(loadings(pc_headlog2), pch = 19, col = sample.colors, main = "Loadings of Samples") 
text(loadings(pc_headlog2)[,1], loadings(pc_headlog2)[,2], labels = colnames(head_50percent_log2), 
       col = sample.colors)
  legend("bottomleft", legend = levels(head_meta$Sample_Type), pch = 16, col = 1:length(levels(head_meta$Sample_Type)),
         y.intersp = 0.7)

```

The distribution of the data is plotted  below before and after imputation
```{r density_head, echo=FALSE}
hist(head_50percent_log2, breaks = 20, xlab = "Log2(head data with the 50% filter)", main = "Histogram of head Data Before Imputation")

```

```{r even_odd_head, echo=FALSE}
#Split data up between the controls and the TBI samples
even_index <- seq(2,34,2) #TBI samples
odd_index <- seq(1,34,2) #controls

#TBI Samples
imputed_head_even <- imputed_head[,even_index]
#Control samples
imputed_head_odd <- imputed_head[,odd_index]
#Reorder data so that first 4 samples are Controls and second set of 4 are the TBI samples
imputed_head_reorder <- imputed_head[,c(odd_index,even_index)]

```

```{r density_imputed_head, echo=FALSE}
hist(imputed_head_reorder, breaks = 20, xlab = "Log2(Imputed head Data with 50% filter", main = "Histogram of Imputed head Data")
```

## Head mFuzz Soft Clustering
Clustering will be implemnted on data set that contains control samples first 17 and TBI samples remaining 17.

First we will look at the variation of the data by calculating the standard deviation of each protein. 
```{r mfuzz_bayesian_headimputed, echo=FALSE}

z = imputed_head_reorder # change accordingly to use in mfuzzy setup funtions

exprValues.s_bayesian <- fuzzyprep_usepreviousImputation(z)
```



Soft clustering is implemented in the function mfuzz using a fuzzy c-means algorithm from e1071 package.
```{r mfuzz_bayesian_head, echo=FALSE}
error <- NA #set error to NA

y = exprValues.s_bayesian #change according to version of data to perform fuzzy clustering
```

To optimize the parameters for fuzzy clustering we must calculate the fuzzier value as well as the number of clusters to divide the data into. The fuzzier "m" and the number of clusters "c" must be chosen in advanced. For "m" we choose a value that prevents clustering of random data. This fuzzy type of clustering is advantageous over hard cluster (e.g. k-means) which commonly detects clusters of random data. 

The function mestimate 
```{r mfuzz_fuzzier_head, echo=FALSE}
m1 <- mestimate(y) 
print(paste0("fuzzier m = ",m1))
```

```{r mfuzz_numberOfClusters_head, include=TRUE}
# Find the error associated with number of cluster 2-15
for(i in 2:20){
  c1 <- mfuzz(y, c=i, m=m1)
  error <- rbind(error, c(i,c1$withinerror))
}
```

Plot error calculated from the number of clusters used.
```{r Error_plot_head, echo=FALSE}
plot(error[,1], error[,2], xlab = "Cluster Number", ylab = "Error")
```

Fuzzy clusters are plotted for 12 clusters.   
*Note first 17 samples are Controls and next set of 17 are TBI samples  
*Note a seed has been set to keep clusters consistent.   
```{r numberOfClusters_beforeMembership_head, echo=FALSE}
c1 <- mfuzz(y, c=12, m=m1)
set.seed(123) #set seed to keep clusters consistent
mfuzz.plot2(y, cl=c1, mfrow = c(3,2), time.labels = c("0",".5", "1", "2","4","6", "8", "12", "16", "24", "48", "72", "168", "336", "504", "672", "840", "0", ".5", "1", "2", "4", "6", "8", "12", "16", "24", "48", "72", "168", "336", "504", "672", "840"), 
            ylab = "Protein Changes", xlab = "Head Time Points (hour)",  col.lab="black", x11=F)

#c("0","0.02", "0.04", "0.08","0.17","0.25", "0.33", "0.50", "0.67", "1.0", "2.0", "3.0", "7.0", "14.0", "21.0", "28.0", "35.0","0.0","0.02", "0.04", "0.08","0.17","0.25", "0.33", "0.50", "0.67", "1.0", "2.0", "3.0", "7.0", "14.0", "21.0", "28.0", "35.0")
```

How many proteins have a membership score greater than 70%, count TRUE.
```{r membershipGreater70_reorder_head, echo=FALSE}
table(rowSums(c1$membership > .70) >0)
```

Heat map of head samples that have a membership value greater than 70%.
Global view of head samples.
```{r heatmap_Highmembership_head, echo=FALSE}
scaleRYG <- colorRampPalette(c("red","black","darkgreen"), space = "rgb")(31)


pheatmap(
  mat               = imputed_head_reorder[rowSums(c1$membership >0.7)>0,],
  color             = scaleRYG,
  border_color      = NA,
  scale = "row",
  cluster_cols = FALSE,
  cluster_rows = TRUE,
  clustering_distance_rows = "euclidean",
  labels_col = c("0","Ctr_.5 hr", "Ctr_1 hr", "Ctr_2 hr","Ctr_4 hr","Ctr_6 hr", "Ctr_8 hr", "Ctr_12 hr", "Ctr_16 hr", "Ctr_24 hr", "Ctr_48 hr", "Ctr_72 hr", "Ctr_168 hr", "Ctr_336 hr", "Ctr_504 hr", "Ctr_672 hr", "Ctr_840 hr", "0",".5 hr", "1 hr", "2 hr","4 hr","6 hr", "8 hr", "12 hr", "16 hr", "24 hr", "48 hr", "72 hr", "168 hr", "336 hr", "504 hr", "672 hr", "840 hr"),
  show_colnames     = TRUE,
  show_rownames     = FALSE,
  #cellwidth = 40,
  #cellheight = .25,
  #annotation_col    = mat_col,
  #annotation_colors = mat_colors,
  drop_levels       = TRUE,
  fontsize          = 8,
  main              = "Log2 Head Data with membership scores greater than 70% (scaled by row)"
)

```

```{r distance_reordered_head,out.width='\\textwidth', fig.height = 8, fig.align='center', include=FALSE}

reorder_hclust <- hclust(dist(imputed_head_reorder[rowSums(c1$membership >0.7)>0,], method = "euclidean"))

clusters <- cutree(reorder_hclust, k = 20)

plot(reorder_hclust, label= FALSE)

#rect.hclust(reorder_hclust, k=20, border = "red")
```

```{r clusters_reordered_head, include=FALSE}
#clusplot(imputed_head_reorder, clusters, lines = 0)
```

## Head mFuzz Soft Clustering - fold change

```{r mfuzz_foldchange_head, include=TRUE}
imputed_head_fold <- imputed_head_reorder[,18:34]-imputed_head_reorder[,1:17]

write.csv(imputed_head_fold, "E:/Projects/Proteomics/DorsophilaHead_Experiment/Imputed_head_foldchange.csv")

# subset features with fold change greater than 1.2 to cluster
important_features_head_foldchange <- imputed_head_fold[rowSums(abs(imputed_head_fold)>1.2)>0,]

# Protein groups with fold change greater than 1.2
table(rowSums(abs(imputed_head_fold)>1.2)>0)

# Protein groups with fold change greater than 2
table(rowSums(abs(imputed_head_fold)>2)>0)

# Protein groups with fold change greater than 4
table(rowSums(abs(imputed_head_fold)>4)>0)

# Protein groups with fold change greater than 8
table(rowSums(abs(imputed_head_fold)>8)>0)


# bin protein groups so that we describe the changes within fold changes of -2 and 2
important_features_foldchange_compress <- important_features_head_foldchange
important_features_foldchange_compress[important_features_foldchange_compress < -2] = -2.1
important_features_foldchange_compress[important_features_foldchange_compress > 2] = 2.1


```

```{r write_csv_imputedhead_foldchange, echo=FALSE}

imputed_headfoldChange_subset <- data.frame(subset,imputed_head_fold)

colnames(imputed_headfoldChange_subset) <- c("ProtienGroup.ID", "Gene.ID","0",".5 hr", "1 hr", "2 hr","4 hr","6 hr", "8 hr", "12 hr", "16 hr", "24 hr", "48 hr", "72 hr", "168 hr", "336 hr", "504 hr", "672 hr", "840 hr")

write.csv(imputed_headfoldChange_subset, "E:/Projects/Proteomics/DorsophilaHead_Experiment/Imputed_head_foldchange.csv")
```

```{r heatmap_important_features_1.2_head, echo=FALSE}
pheatmap(
  mat               = important_features_head_foldchange, #[rowSums(c2$membership >0.5)>0,],
  color             = scaleRYG,
  border_color      = NA,
  #scale = "row",
  cluster_cols = FALSE,
  cluster_rows = TRUE,
  labels_col = c("0",".5 hr", "1 hr", "2 hr","4 hr","6 hr", "8 hr", "12 hr", "16 hr", "24 hr", "48 hr", "72 hr", "168 hr", "336 hr", "504 hr", "672 hr", "840 hr"),
  #breaks = breaksList,
  clustering_distance_rows = "euclidean",
  show_colnames     = TRUE,
  show_rownames     = FALSE,
  #cellwidth = 40,
  #cellheight = .25,
  #annotation_col    = mat_col,
  #annotation_colors = mat_colors,
  drop_levels       = TRUE,
  fontsize          = 10,
  main              = "Heat map of head protein with fold change < -1.2, > 1.2"
)

pheatmap(
  mat               = important_features_foldchange_compress, #[rowSums(c2$membership >0.5)>0,],
  color             = scaleRYG,
  border_color      = NA,
  #scale = "row",
  cluster_cols = FALSE,
  cluster_rows = TRUE,
  labels_col = c("0",".5 hr", "1 hr", "2 hr","4 hr","6 hr", "8 hr", "12 hr", "16 hr", "24 hr", "48 hr", "72 hr", "168 hr", "336 hr", "504 hr", "672 hr", "840 hr"),
  #breaks = breaksList,
  clustering_distance_rows = "euclidean",
  show_colnames     = TRUE,
  show_rownames     = FALSE,
  #cellwidth = 40,
  #cellheight = .25,
  #annotation_col    = mat_col,
  #annotation_colors = mat_colors,
  drop_levels       = TRUE,
  fontsize          = 10,
  main              = "Fold Change with compressed Head Data, binning <-2 and >2"
)


```

```{r mfuzz_clustering_head, include=TRUE}
fold_hclust <- hclust(dist(important_features_foldchange_compress, method = "euclidean"))
clusters <- cutree(fold_hclust, k = 6) #k = 20)

clusplot(important_features_foldchange_compress, clusters, lines = 0)

plot(fold_hclust, label= FALSE)
rect.hclust(fold_hclust, k=6, border = "red")
```

# Fold Change Head Data Fuzzy Clustering
```{r mfuzz_bayesian_imputed_foldchange_head, include=TRUE}

#start with the re-ordered hemo data
z = important_features_foldchange_compress # change accordingly to use in mfuzzy setup funtions

exprValues.s_bayesian <- fuzzyprep_usepreviousImputation_foldchange(z)
```

```{r mfuzz_bayesian_foldchange_head, include=TRUE}
error <- NA #set error to NA

y = exprValues.s_bayesian #change according to version of data to perform fuzzy clustering
```

The function mestimate 
 
```{r mfuzz_bayesian_imputed_foldchange_error_, include=TRUE}
m1 <- mestimate(y) 

# 
for(i in 2:20){
  c1 <- mfuzz(y, c=i, m=m1)
  error <- rbind(error, c(i,c1$withinerror))
}
```

Plot error calculated from imputed hemo data set that has been reordered. 
```{r error_head_foldChange, echo=FALSE}
plot(error[,1], error[,2])
```

```{r numberOfClusters_foldchange_head, echo=FALSE}
c1 <- mfuzz(y, c=12, m=m1)
set.seed(123)
mfuzz.plot2(y, cl=c1, mfrow = c(3,2), time.labels = c("0",".5", "1", "2","4","6", "8", "12", "16", "24", "48", "72", "168", "336", "504", "672", "840"),
            ylab = "Protein Changes", xlab = "Head Time Points",  col.lab="black", x11=F)
mfuzz.plot2(y, cl=c1, mfrow = c(3,2), time.labels = c("0",".5", "1", "2","4","6", "8", "12", "16", "24", "48", "72", "168", "336", "504", "672", "840"), 
            min.mem = 0.50, ylim.set = c(-2,2), ylab = "Protein Changes", xlab = "Head Time Points",  col.lab="black", x11=F)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
